### 2024-08-10 machine-learning.qmd
20240809 - updates

nfl_expectedFantasyPoints_weekly <- nfl_expectedFantasyPoints_weekly %>% 
  mutate(
    receptions = as.integer(receptions))

@@

nfl_advancedStatsPFR_seasonal %>% 
  group_by(gsis_id, season, team) %>% 
  filter(n() > 1, !is.na(gsis_id)) %>% 
  select(gsis_id, pfr_id, season, team, everything()) %>% 
  head()
```

@@

playerSeasonListToMerge <- list(
  player_stats_seasonal_offense %>% filter(!is.na(gsis_id), !is.na(season), !is.na(team))
)

playerSeasonTeamListToMerge <- list(
  nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season), !is.na(team))
)

@@

playerSeasonWeekPositionListToMerge <- list(
  nfl_depthCharts %>% filter(!is.na(gsis_id), !is.na(season), !is.na(week)),
)

@@

playerSeasonTeamMerged <- playerSeasonTeamListToMerge %>% 
  reduce(
    powerjoin::power_full_join,
    by = c("gsis_id","season"),
    conflict = coalesce_xy)

@@

seasonalData <- powerjoin::power_full_join(
  idSeasonMerged,
  idMerged,
  by = "gsis_id",
  conflict = coalesce_xy
)

seasonalAndWeeklyData <- powerjoin::power_full_join(
  idSeasonWeekMerged,
  seasonalData,
  by = c("gsis_id","season"),
  conflict = coalesce_xy
)

# check on duplicate id-by-season data (nfl_rosters, nfl_advancedStatsPFR_seasonal)
seasonalData %>% 
  select(gsis_id, season)

seasonalData %>% 
  select(gsis_id, season) %>% 
  unique()
```

@@

seasonalData_subset <- seasonalData %>% 
  dplyr::select(-c(
    birth_date, loaded, full_name, player_name, player_display_name, display_name, suffix, headshot_url, player, pos,
    espn_id, sportradar_id, yahoo_id, rotowire_id, pff_id, fantasy_data_id, sleeper_id, pfr_id,
    pfr_player_id, cfb_player_id, pfr_player_name, esb_id, gsis_it_id, smart_id,
    college, college_name, team_abbr, current_team_id, college_conference, draft_club, status_description_abbr,
    status_short_description, short_name, headshot, uniform_number, jersey_number, first_name, last_name,
    football_name, team
    ))
```

@@

@@

seasonalData_subset_imp <- missForest::missForest(
  seasonalData_subset %>% select(ageCentered20, ageCentered20Quadratic, fantasy_points) %>% data.frame,
  verbose = TRUE,
  mtry = 100,
  parallelize = "forests")
```

@@

```{r}
smerf <- LongituRF::MERF(
  X = seasonalData[,c("ageCentered20")] %>% as.matrix(),
  Y = seasonalData$fantasy_points,
  Z = seasonalData[,c("ageCentered20")] %>% as.matrix(),
  id = seasonalData$gsis_id,
  time = seasonalData$ageCentered20,
  mtry = 2,
  ntree = 500,
  sto = "BM")

>>>>

nfl_expectedFantasyPoints_weekly <- nfl_expectedFantasyPoints_weekly %>% 
  mutate(
    season = as.integer(season),
    receptions = as.integer(receptions))

@@

nfl_advancedStatsPFR_seasonal %>% 
  group_by(gsis_id, season) %>% 
  filter(n() > 1, !is.na(gsis_id)) %>% 
  select(gsis_id, pfr_id, season, team, everything()) %>% 
  head()
```

@@

playerSeasonListToMerge <- list(
  player_stats_seasonal_offense %>% filter(!is.na(gsis_id), !is.na(season)),
  nfl_advancedStatsPFR_seasonal %>% filter(!is.na(gsis_id), !is.na(season))
)

@@

playerSeasonWeekPositionListToMerge <- list(
  nfl_depthCharts %>% filter(!is.na(gsis_id), !is.na(season), !is.na(week))
)

@@

@@

seasonalData <- powerjoin::power_full_join(
  playerSeasonMerged,
  playerMerged,
  by = "gsis_id",
  conflict = coalesce_xy
)

seasonalAndWeeklyData <- powerjoin::power_full_join(
  playerSeasonWeekMerged,
  seasonalData,
  by = c("gsis_id","season"),
  conflict = coalesce_xy
)

# check on duplicate id-by-season data (nfl_rosters, nfl_advancedStatsPFR_seasonal)

seasonalData %>% 
  group_by(gsis_id, season) %>% 
  filter(n() > 1) %>% 
  head()
```

@@

seasonalData_subset <- seasonalData %>% 
  dplyr::select(-any_of(c(
    "birth_date", "loaded", "full_name", "player_name", "player_display_name", "display_name", "suffix", "headshot_url", "player", "pos",
    "espn_id", "sportradar_id", "yahoo_id", "rotowire_id", "pff_id", "fantasy_data_id", "sleeper_id", "pfr_id",
    "pfr_player_id", "cfb_player_id", "pfr_player_name", "esb_id", "gsis_it_id", "smart_id",
    "college", "college_name", "team_abbr", "current_team_id", "college_conference", "draft_club", "status_description_abbr",
    "status_short_description", "short_name", "headshot", "uniform_number", "jersey_number", "first_name", "last_name",
    "football_name", "team"
  ))) %>% 
  filter(!is.na(season))
```

@@

### Lag Fantasy Points {#sec-lagFantasyPoints}

```{r}
seasonalData_subset <- seasonalData_subset %>% 
  arrange(gsis_id, season) %>% 
  group_by(gsis_id) %>% 
  mutate(
    fantasyPoints_lag = lead(fantasy_points)
  ) %>% 
  ungroup()
```

@@

seasonalData_subset_imp <- missForest::missForest(
  seasonalData_subset %>% select(ageCentered20, ageCentered20Quadratic, fantasyPoints_lag) %>% data.frame,
  verbose = TRUE,
  mtry = 100,
  parallelize = "forests")
```

@@

```{r}
smerf <- LongituRF::MERF(
  X = seasonalData_subset_imp$ximp[,c("ageCentered20")] %>% as.matrix(),
  Y = seasonalData_subset_imp$ximp[,c("fantasyPoints_lag")],
  Z = seasonalData_subset_imp$ximp[,c("ageCentered20")] %>% as.matrix(),
  id = seasonalData_subset$gsis_id,
  time = seasonalData_subset_imp$ximp[,c("ageCentered20")] %>% as.matrix(),
  mtry = 2,
  ntree = 500,
  sto = "BM")
